Alone Together: Compositional Reasoning and Inference for Weak Isolation

### 2017-10-18 inference.tex
Fixing typos.
Minor Changes in RG-rules

The correspondence between the inference rules and the RG judgment of
\S~\ref{sec:reasoning} is stated thus\footnote{The proof can be found in
the supplementary material.}:
\begin{theorem}
\label{thm:inference-sound}
  \emph{Forall} i,$R$,$I$,$c$,$\F$, if $\stable(\R,I)$ and $c \elabsto \F$,
  then:\\\vspace*{-0.2cm}
  \begin{smathpar}
  \begin{array}{c}
  \R \vdash \hoare{\lambda(\stl,\stg).~\stl=\emptyset \conj
  I(\stg)}{c}{\lambda(\stl,\stg).\stl \subseteq \F(\emptyset,\stg)}
  \end{array}
  \end{smathpar}
\end{theorem}
\noindent where the set expression $\F(\emptyset,\stg)$ has a natural
interpretation as discussed below.
% Where, ${\sf true} = \lambda(\stl,\stg).~true$ and $\stable(\R,I)
% \Leftrightarrow \forall\stl,\stg,\stg'. I(\stg) \conj
% \R(\stl,\stg,\stg') \Rightarrow \I(\stg')$.

@@

Theorem~\ref{thm:inference-sound} lets us replace the local judgment
of the \rulelabel{RG-Txn} rule (Fig.~\ref{fig:rg-rules}) by a state
transformer inference judgment. The soundness of a transaction's guarantee can now
be established w.r.t the effect $\F$ of the body. The
\rulelabel{RG-Txn} rule so updated is shown
below:
\begin{smathpar}
\begin{array}{c}
\RULE
{
  \stable(R,\I)\spc
  \stable(R,I)\\
   \R_e(\stl,\stg,\stg') \Leftrightarrow \exists \stg_1.  R(\stg, \stg') \wedge \I_e(\stl, \stg_1, \stg') \spc  c \elabsto \F\\
  \R_c(\stl,\stg,\stg') \Leftrightarrow \exists \stg_1.  R(\stg, \stg') \wedge \I_c(\stl, \stg_1, \stg') \spc \stable(\R_c,\F)\\
  \forall \stg.~ G(\stg, F(\emptyset,\stg) \gg \stg)\spc
  \forall \stg,\stg'.~I(\stg) \wedge G(\stg,\stg') \Rightarrow I(\stg')\\
}
{
  \rg{I,R}{\ctxn{i}{\I}{c}}{G,I}
}
\end{array}
\end{smathpar}
Automating the application of the \rulelabel{RG-Txn} rule for a
transaction requires automating the multiple implication checks in
the premise. While $R$, $\R$, $\R_c$, $\I$ and $I$ are formulas in
first-order logic (FOL) with a relatively simple structure, $\F$
is an expression in the set language $\SL$
(Fig.~\ref{fig:logic-syntax}) with a possibly complex structure.
Fortunately, however, there exists a semantics-preserving translation
from $\SL$ to a restricted subset of first-order logic (FOL) that
lends itself to automatic reasoning.

>>>>

The correspondence between the inference rules and the RG judgment of
\S~\ref{sec:reasoning} is stated thus\footnote{The proof can be found in
the supplementary material.}:
\begin{theorem}
\label{thm:inference-sound}
  \emph{Forall} i,$R$,$I$,$c$,$\F$, if $\stable(\R,I)$ and $c \elabsto \F$,
  then:\\\vspace*{-0.2cm}
  \begin{smathpar}
  \begin{array}{c}
  \R \vdash \hoare{\lambda(\stl,\stg).~\stl=\emptyset \conj
  I(\stg)}{c}{\lambda(\stl,\stg).\stl = \F(\emptyset,\stg)}
  \end{array}
  \end{smathpar}
\end{theorem}
\noindent where the set expression $\F(\emptyset,\stg)$ has a natural
interpretation as discussed below.

\begin{proof}
We only provide a brief proof sketch here. We prove a slightly stronger statement :

\emph{Forall} i,$R$,$I$,$c$,$\F$,$s$ if $\stable(\R,I)$ and $c \elabsto \F$,
  then:\\

\begin{smathpar}
  \begin{array}{c}
  \R \vdash \hoare{\lambda(\stl,\stg).~\stl=s \conj
  I(\stg)}{c}{\lambda(\stl,\stg).\stl = \F(s,\stg)}
  \end{array}
  \end{smathpar}

We use structural induction on $c$ to prove the statement. Assume that $P \Leftrightarrow \lambda(\stl,\stg).~\stl=s \conj I(\stg)$ and $Q \Leftrightarrow \lambda(\stl,\stg).\stl = \F(s,\stg)$. For each type of $c$, our strategy is to show the premises of the corresponding RG-rule. The base cases correspond to the INSERT, UPDATE and DELETE statements. In all these cases, the form of $F$ directly corresponds to the condition linking the pre-condition $P$ and post-condition $Q$ in the premises of the RG-rules. If $\stable(\R, F)$, then $\stabilize{F} = F$ and hence the proof directly follows. Otherwise, $\stabilize{F} = \lambda (\stl, \stg).\existsl(\stg', I(\stg'), F(\stl, \stg'))$. Since the pre-condition $P(\stl, \stg)$ asserts $I(\stg)$, any state which obeys $P$ provides the existential $\stg'$ in the above set expression, and hence $F$ applied on that state matches the condition linking $P$ and $Q$. The stability of $P$ with respect to $\R$ follows directly from the stability of $I$.

The proofs for SELECT, SEQ and IF-THE-ELSE directly use the inductive hypothesis to infer the RG-judgements present in the premises of their corresponding RG-rules. The interesting case is the FOREACH statement, for which we use the loop invariant $\psi(\stl, \stg) \Leftrightarrow \stl = s \cup y \bind (\lambda z.~\F(s, \stg))$, (where assuming that $c$ is the body of the loop, $c \elabsto F$) to prove all the premises of \rulelabel{RG-Foreach}. Using the same notation as the rule $\rulelabel{RG-Foreach}$, $y$ refers to the records already processed in previous iterations of the loop, while $z$ refers to the record being processed in the current iteration. At the beginning of the loop $[\phi/y]\psi(\stl, \stg)$ just reduces to $\stl = s$ which is implied by the pre-condition $P$. From the inductive hypothesis, we can infer that each iteration corresponds to the application of $F$. Since all iterations are assumed to be independent of each other, and $z$ is bound to a record in $x$ for each iteration, we can infer that at the end of every iteration the loop invariant $[y \cup \{z\}/y]\psi$ will be satisfied. 
\end{proof}
% Where, ${\sf true} = \lambda(\stl,\stg).~true$ and $\stable(\R,I)
% \Leftrightarrow \forall\stl,\stg,\stg'. I(\stg) \conj
% \R(\stl,\stg,\stg') \Rightarrow \I(\stg')$.

@@

Theorem~\ref{thm:inference-sound} lets us replace the local judgment
of the \rulelabel{RG-Txn} rule (Fig.~\ref{fig:rg-rules}) by a state
transformer inference judgment. The soundness of a transaction's guarantee can now
be established w.r.t the effect $\F$ of the body. The
\rulelabel{RG-Txn} rule so updated is shown
below:
\begin{smathpar}
\begin{array}{c}
\RULE
{
  \stable(R,\I)\spc
  \stable(R,I)\spc
   \R_e(\stl,\stg,\stg') \Leftrightarrow \exists \stg_1.  R(\stg, \stg') \wedge \I_e(\stl, \stg_1, \stg') \spc  c \elabsto \F\\
  \R_c(\stl,\stg,\stg') \Leftrightarrow \exists \stg_1.  R(\stg, \stg') \wedge \I_c(\stl, \stg_1, \stg') \spc \stable(\R_c,\F)\\
  \forall \stl,\stg.~ \stl \in F(\emptyset,\stg) \Rightarrow 
    G(\stg, \stl\gg\stg)\spc
  \forall \stg,\stg'.~I(\stg) \wedge G(\stg,\stg') \Rightarrow I(\stg')\\
}
{
  \rg{I,R}{\ctxn{i}{\I}{c}}{G,I}
}
\end{array}
\end{smathpar}
Automating the application of the \rulelabel{RG-Txn} rule for a
transaction requires automating the multiple implication checks in
the premise. While $R$, $\R$, $\R_c$, $\I$ and $I$ are formulas in
first-order logic (FOL) with a relatively simple structure, $\F$
is an expression in the set language $\SL$
(Fig.~\ref{fig:logic-syntax}) with a possibly complex structure.
Fortunately, however, there exists a semantics-preserving translation
from $\SL$ to a restricted subset of first-order logic (FOL) that
lends itself to automatic reasoning.

### 2017-10-18 reasoning.tex
Minor Changes in RG-rules

A central feature of a rely-guarantee judgment is a stability
condition that requires the validity of an assertion $\phi$ to be
unaffected by the interference, i.e., the rely relation $R$. In
conventional RG, stability is defined as follows, where $\sigma$ denotes a
state:
\begin{smathpar}
\begin{array}{lcl}
\stable(R,\phi) & \Leftrightarrow & \forall \sigma,\sigma'.~
\phi(\sigma) \conj R(\sigma,\sigma') \Rightarrow \phi(\sigma')\\
\end{array}
\end{smathpar}
Due to the presence of local and global database states, and the
availability of an isolation specification, we use multiple
definitions of stability in Fig.~\ref{fig:rg-rules}, but they all
convey the same intuition as above.

\rulelabel{RG-Txn} is the top-level rule that lets us prove a
transaction preserves the high-level invariant $I$ when executed under
the required isolation as specified by $\I$. It depends on a
transaction-local judgment to verify the body ($c$) of a transaction with
id $i$. The precondition $P$ of $c$ must follow from the fact that the
transaction-local database ($\stl$) is initially empty, and the global
database satisfies the high-level invariant $I$. The rely relation
($\R_e$) for the local judgment is a ternary relation computed as $R$
modulo $\I_e$:
\begin{smathpar}
\begin{array}{lcl}
%\R_e(\stl,\stg,\stg') & \Leftrightarrow & R(\stg,\stg') \conj
%\I_e(\stl,\stg,\stg')
  \R_e(\stl,\stg,\stg') & \Leftrightarrow & \exists \stg_1.  R(\stg, \stg') \wedge \I_e(\stl, \stg_1, \stg')
\end{array}
\end{smathpar}
\noindent While $R$ describes changes in the global database state, $\I_e$
constrains the global effects visible to the transaction.
Recall however that $\I_e$ uses the global state in effect at the
point the transaction last took a step ($\stg_1$) to determine whether
the transaction can proceed from the current global state ($\stg$).
Moreover, the global rely relation $R$ defines the permissible
interferences from $\Delta$ to a future global state ($\Delta'$).  The
local rely relation $\R_e$ determines allowed interferences as
specified by $\I_e$ assuming $\stg_1$.  In effect, the existential on
$\stg_1$ defines this rely relation to be an over-approximation on
the interferences in the presence of which the transaction can
proceed, simplifying verification burden.

%Often, this significantly simplifies the form of $\R_l$ irrespective
%of $R$. For example, when a transaction is executed against a
%snapshot of the database (i.e. $\mathbb{I}_{ss}$), $\R_l$ will be the
%identity function, since if $\stg \neq \stg'$, there does not exist
%any $\stg_1$ which satisfies the constraint for
%$\R_l(\stl,\stg,\stg')$.

@@

Recall that $P$ and $Q$ of the transaction-local RG judgment are
binary assertions; they relate local and global database states. The
local judgment rules require one or both of them to be stable. The
stability of a binary assertion $Q$ w.r.t a ternary rely relation $\R$
is defined as:
\begin{smathpar}
\begin{array}{c}
\forall \stl,\stg,\stg'.~ Q(\stl,\stg) \conj \R(\stl,\stg,\stg')
\Rightarrow Q(\stl,\stg')
\end{array}
\end{smathpar}
That is, if $Q$ relates $\stl$ to $\stg$, and an interference allowed
by the isolation specification (which implicitly considers the local
state $\stl$) takes $\stg$ to $\stg'$, then $Q$ must also relate $\stl$
to $\stg'$.

@@

A characteristic of RG reasoning is that stability of an assertion is
always proven w.r.t to $R$, and not $R^{*}$, although interference may
include multiple environment steps, and $R$ only captures a single
step. This is nonetheless sound due to inductive reasoning: if
$Q$ is preserved by every step of $R$, then $Q$ is preserved by
$R^{*}$, and vice-versa.  However, such reasoning does not extend
naturally to isolation-constrained interference because $R^{*}$ modulo
$\I$ is not same as $\R^{*}$; the former is a transitive relation
constrained by $\I$, whereas the latter is the transitive closure of an
$\I$-constrained relation. We therefore introduce a side-condition on
$\I$ that restores the equality. The condition requires $\I$ to allow
an interference $R^{*}(\stg,\stg'')$, for two database states $\stg$
and $\stg''$, only if it also allows interference for every prefix of
$R^{*}(\stg,\stg'')$. In other words, if $\I$ disallows interference
from $\stg$ to $\stg'$, then an $R$-step from $\stg'$ to $\stg''$
should not make the interference from $\stg$ to $\stg''$ valid.  This
stability condition on $\I$ is defined formally thus:
\begin{smathpar}
\begin{array}{lcl}
  \stable(R,\I) & \Leftrightarrow & \forall \stl,\stg,\stg',\stg''.~
  \neg\I(\stl,\stg,\stg') \conj R(\stg',\stg'') \Rightarrow
  \neg\I(\stl,\stg,\stg'')
\end{array}
\end{smathpar}
It can be easily verified that the above stability condition is
satisfied by the isolation axioms from Sec.~\ref{sec:isolation}. For
instance, $\I_{ss}$, the snapshot axiom, is stable because if
$\I_{ss}$ is invalid ($\neg\I_{ss}$), then an interference has already
modified a record, and no further interference will restore the
original record, because the original record bears the id of a
transaction that has already committed.

@@

>>>>

A central feature of a rely-guarantee judgment is a stability
condition that requires the validity of an assertion $\phi$ to be
unaffected by the interference, i.e., the rely relation $R$. In
conventional RG, stability is defined as follows, where $\sigma$ denotes a
state:
\begin{smathpar}
\begin{array}{lcl}
\stable(R,\phi) & \Leftrightarrow & \forall \sigma,\sigma'.~
\phi(\sigma) \conj R(\sigma,\sigma') \Rightarrow \phi(\sigma')\\
\end{array}
\end{smathpar}
Due to the presence of local and global database states, and the
availability of an isolation specification, we use multiple
definitions of stability in Fig.~\ref{fig:rg-rules}, but they all
convey the same intuition as above. In our setting, we only need to prove the stability of an assertion ($Q$) against those environment steps which lead to a global database state on which the transaction itself can take its next step according to its isolation specification ($\I$).

\begin{smathpar}
\begin{array}{lcl}
\stable(R, Q) & \Leftrightarrow & \forall \stl, \stg, \stg'. Q(\stl, \stg) \wedge R^{*}(\stg, \stg') \wedge \I(\stl, \stg, \stg') \Rightarrow Q(\stl, \stg')
\end{array}
\end{smathpar}

The above condition uses the transitive closure of $R$, while the standard RG stability assertion only uses $R$ and then relies on an inductive argument to prove stability for the transitive closure. We cannot directly replace $R^{*}$ by $R$ in the above condition because the inductive argument would not hold, as $R^{*}$ modulo $\I$ is not the same as ($R$ modulo $\I$$)^{*}$.To obtain an equivalent form in our setting, we require an additional condition on the isolation specification, which we call the stability condition on $\I$. 
The condition requires $\I$ to allow
an interference $R^{*}(\stg,\stg'')$, for two database states $\stg$
and $\stg''$, only if it also allows interference for every prefix of
$R^{*}(\stg,\stg'')$. In other words, if $\I$ disallows interference
from $\stg$ to $\stg'$, then an $R$-step from $\stg'$ to $\stg''$
should not make the interference from $\stg$ to $\stg''$ valid.  This
stability condition on $\I$ is defined formally thus:
\begin{smathpar}
\begin{array}{lcl}
  \stable(R,\I) & \Leftrightarrow & \forall \stl,\stg,\stg',\stg''.~
  \neg\I(\stl,\stg,\stg') \conj R(\stg',\stg'') \Rightarrow
  \neg\I(\stl,\stg,\stg'')
\end{array}
\end{smathpar}
It can be easily verified that the above stability condition is
satisfied by the isolation axioms from Sec.~\ref{sec:isolation}. For
instance, $\I_{ss}$, the snapshot axiom, is stable because if
$\I_{ss}$ is invalid ($\neg\I_{ss}$), then an interference has already
modified a record, and no further interference will restore the
original record, because the original record bears the id of a
transaction that has already committed. We can now generalize the stability condition for any arbitrary assertion $Q$ by only considering those environment steps which in addition to preserving $Q$, also preserve the condition that the transaction can continue to take its next step, i.e. the condition $\I$ holds on both the beginning and the end states.

 \begin{smathpar}
\forall \stl, \stg, \stg', \stg_1. Q(\stl, \stg) \wedge \I(\stl, \stg_1, \stg) \wedge R^{*}(\stg, \stg') \wedge \I(\stl, \stg_1, \stg') \Rightarrow Q(\stl, \stg')
\end{smathpar}

In the above condition, we use a separate notation ($\stg_1$) for the global database state that the transaction witnessed during its last step. $Q$ does not constrain this global database state, but the stability of $\I$ ensures that any database state constrained by $Q$ must also obey $\I$. We can now replace the transitive closure in the above condition with the single step relation $R$, backed by the inductive argument for stability. Pushing the universal quantifier for $\stg_1$ inside the antecedent and separating the predicates involving $\stg_1$, we can rewrite the condition in the form below, by defining a new ternary relation $\R$, which can be thought of as $R$ modulo $\I$:%stating that an assertion $Q$ on a database state $\stg$ on which the transaction can take its next step (hence the condition $\I(\stl, \stg_1, \stg)$) is stable w.r.t. environment steps to $\stg'$

\begin{smathpar}
\begin{array}{lcl}
\stable((R \backslash \I), Q) & \Leftrightarrow & \forall \stl,\stg,\stg'. Q(\stl,\stg) \wedge (R \backslash \I)(\stl,\stg,\stg') \Rightarrow Q(\stl,\stg')\\
(R \backslash \I)(\stl, \stg, \stg') & \Leftrightarrow & \exists \stg_1. R(\stg, \stg') \wedge \I(\stl, \stg_1, \stg) \wedge \I(\stl, \stg_1, \stg')
\end{array}
\end{smathpar}

The above condition often significantly simplifies the form of $\R$ irrespective
of $R$. For example, when a transaction is executed against a
snapshot of the database (i.e. $\mathbb{I}_{ss}$), $\R$ will be the
identity function, since if $\stg \neq \stg'$, there does not exist
any $\stg_1$ which satisfies the constraint for
$\R_l(\stl,\stg,\stg')$.

\rulelabel{RG-Txn} is the top-level rule that lets us prove a
transaction preserves the high-level invariant $I$ when executed under
the required isolation as specified by $\I$. It depends on a
transaction-local judgment to verify the body ($c$) of a transaction with
id $i$. The precondition $P$ of $c$ must follow from the fact that the
transaction-local database ($\stl$) is initially empty, and the global
database satisfies the high-level invariant $I$. The rely relation
($\R_e$) is obtained from the global rely relation $R$ and the isolation specification $\I_e$ as explained above. Note that $\I_e$ constrains the global effects visible to the transaction while it is executing but has not yet committed.

%for the local judgment is a ternary relation computed as $R$
%modulo $\I_e$:
%\begin{smathpar}
%\begin{array}{lcl}
%\R_e(\stl,\stg,\stg') & \Leftrightarrow & R(\stg,\stg') \conj
%\I_e(\stl,\stg,\stg')
%  \R_e(\stl,\stg,\stg') & \Leftrightarrow & \exists \stg_1.  R(\stg, \stg') \wedge \I_e(\stl, \stg_1, \stg) \wedge \I_e(\stl, \stg_1, \stg')
%\end{array}
%\end{smathpar}
%\noindent While $R$ describes changes in the global database state, $\I_e$
%constrains the global effects visible to the transaction.
%Recall however that $\I_e$ uses the global state in effect at the
%point the transaction last took a step ($\stg_1$) to determine whether
%the transaction can proceed from the current global state ($\stg$).
%Moreover, the global rely relation $R$ defines the permissible
%interferences from $\Delta$ to a future global state ($\Delta'$).  The
%local rely relation $\R_e$ determines allowed interferences as
%specified by $\I_e$ assuming $\stg_1$.  In effect, the existential on
%$\stg_1$ defines this rely relation to be an over-approximation on
%the interferences in the presence of which the transaction can
%proceed, simplifying verification burden.

@@

Recall that $P$ and $Q$ of the transaction-local RG judgment are
binary assertions; they relate local and global database states. The
local judgment rules require one or both of them to be stable with respect to the constrained rely relation $\R_e$.

%The
%stability of a binary assertion $Q$ w.r.t a ternary rely relation $\R$
%is defined as:
%\begin{smathpar}
%\begin{array}{c}
%\forall \stl,\stg,\stg'.~ Q(\stl,\stg) \conj \R(\stl,\stg,\stg')
%\Rightarrow Q(\stl,\stg')
%\end{array}
%\end{smathpar}
%That is, if $Q$ relates $\stl$ to $\stg$, and an interference allowed
%by the isolation specification (which implicitly considers the local
%state $\stl$) takes $\stg$ to $\stg'$, then $Q$ must also relate $\stl$
%to $\stg'$.

@@

%A characteristic of RG reasoning is that stability of an assertion is
%always proven w.r.t to $R$, and not $R^{*}$, although interference may
%include multiple environment steps, and $R$ only captures a single
%step. This is nonetheless sound due to inductive reasoning: if
%$Q$ is preserved by every step of $R$, then $Q$ is preserved by
%$R^{*}$, and vice-versa.  However, such reasoning does not extend
%naturally to isolation-constrained interference because $R^{*}$ modulo
%$\I$ is not same as $\R^{*}$; the former is a transitive relation
%constrained by $\I$, whereas the latter is the transitive closure of an
%$\I$-constrained relation. We therefore introduce a side-condition on
%$\I$ that restores the equality. The condition requires $\I$ to allow
%an interference $R^{*}(\stg,\stg'')$, for two database states $\stg$
%and $\stg''$, only if it also allows interference for every prefix of
%$R^{*}(\stg,\stg'')$. In other words, if $\I$ disallows interference
%from $\stg$ to $\stg'$, then an $R$-step from $\stg'$ to $\stg''$
%should not make the interference from $\stg$ to $\stg''$ valid.  This
%stability condition on $\I$ is defined formally thus:
%\begin{smathpar}
%\begin{array}{lcl}
%  \stable(R,\I) & \Leftrightarrow & \forall \stl,\stg,\stg',\stg''.~
%  \neg\I(\stl,\stg,\stg') \conj R(\stg',\stg'') \Rightarrow
%  \neg\I(\stl,\stg,\stg'')
%\end{array}
%\end{smathpar}
%It can be easily verified that the above stability condition is
%satisfied by the isolation axioms from Sec.~\ref{sec:isolation}. For
%instance, $\I_{ss}$, the snapshot axiom, is stable because if
%$\I_{ss}$ is invalid ($\neg\I_{ss}$), then an interference has already
%modified a record, and no further interference will restore the
%original record, because the original record bears the id of a
%transaction that has already committed.

@@

\begin{proof}
Here, we provide only a brief proof sketch. The most important rule is the top-level rule \rulelabel{RG-Txn}, which proves that a transaction $c$ which begins its execution in global database state satisfying $I$ and suffering interference $R$ while executing under isolation specification $\I$ finishes its execution in a database state also satisfying $I$, and also guarantees that its commit step satisfies $G$. The rule uses the transaction-local RG judgement $\R_e \vdash \rg{P}{c}{Q}$. By \rulelabel{E-Txn-Start}, the local and global database states at the start of a transaction satisfy $P$, and the only challenge is that environment steps in an execution covered by $\R_e \vdash \rg{P}{c}{Q}$ are in $\R_e$, while the top-level judgement requires environment steps in $R$. We show that it is enough to consider only those enviroment steps in $\R_e$. First, we use an inductive argument and stability of $\I_e$ ($\stable(R, \I_e)$) to show that any execution in which the transaction does not get stuck and completes all its steps must always preserve the isolation specification $\I_e$ after every environment step. Intuitively, this is because once $\I_e$ gets broken after some environment step, it will continue to remain broken and the transaction would get stuck. Then, we show that $\R$ contains exactly those environment steps which preserve $\I$. Hence, the local-level RG judgement can be soundly used, which guarantees that after the transaction finishes its execution, its local state $\stl$ and global state $\stg$ will satisfy the invariant $Q$. Environment steps between the last step of the transaction and its commit step can modify the global state, and hence we also require $Q$ to be stable againt $R$. Again, we use an inductive argument, the stability of $\I_c$, and the fact that the transaction does not get stuck and must execute its commit step to show that all environment steps must preserve $\I_c$, and hence it is enough to require $\stable(\R_c, Q)$. $Q$ guarantees that the commit step is in $G$, and $G$ in turn guarantees that after execution, the global database state will obey the invariant $I$.

The local RG-judgement rules are mostly straightforward and directly follow from the operational semantics. At this point, the transaction can only change its local database state by adding new records to it, and hence all the rules have the form : $P(\stl, \stg)$ and $\stl'$ is the new state should satisfy $Q(\stl', \stg)$. An important thing to note is that the rules for Update, Select, Delete and Insert only require the pre-condition $P$ to be stable against interference. This is because the local RG-judgement rules only make assertions about the database state just after the transaction takes its last step. For the aforementioned commands, the last step would be the only step taken by the transaction, and hence there would be no environment steps which could affect $Q$ (the environment steps between the last execution step of the transaction and the commit step would be controlled by $\I_c$ and would be taken care of by $\stable(\R_c, Q)$ at the top level).

The rule $\rulelabel{RG-Foreach}$ makes use of an inductive loop invariant $\psi$ which assumes that the variable $y$ is bound to records which have been processed in previous iterations and $z$ is bound to the record processed in the current iteration. Since the operational semantics ensure that every iteration processes a unique record, the proof is straightforward.  
\end{proof}
